Yeast RNA-seq 48 sample simulation study: Unimodal alternative, 2000 DE genes

### 2018-09-13 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Refine informative covariate in yeast
Make strongly informative covariate stronger and weakly informative
covariate weaker.

# results files that will be generated
resfile_d5 <- file.path(resdir, "yeastIIH-results-de5.rds")
resfile_d10 <- file.path(resdir, "yeastIIH-results-de10.rds")
resfile_d5_w <- file.path(resdir, "yeastIIHW-results-de5.rds")
resfile_d10_w <- file.path(resdir, "yeastIIHW-results-de10.rds")

resfile_d5_uninfCov <- file.path(resdir, "yeastIIH-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "yeastIIH-results-de10-uninfCov.rds")
resfile_d5_uninfCov_w <- file.path(resdir, "yeastIIHW-results-de5-uninfCov.rds")
resfile_d10_uninfCov_w <- file.path(resdir, "yeastIIHW-results-de10-uninfCov.rds")

@@

  covar <- abs(rnorm(nrow(dds)))

@@

  if(nDE > 0){
    # pick random set of nDE genes to add signal to
    if(strongCovariate){
      wts <- covar
    }else{
      wts <- abs(covar + rnorm(length(covar)))
    }
    DE <- sample(1:nrow(dds_test), nDE, prob = wts)
    truth[DE] <- TRUE

@@

  if (uninformativeCovariate)
    geneExp <- mutate(geneExp, ind_covariate = abs(rnorm(length(covar))))

@@

# Sim D5: DE Comparison 5v5

@@

## Generate a list of SB results objects

@@

## Plot average results over replications

@@

# Sim D10: DE Comparison 10v10

@@

# Sim D5: 5v5 Weak Covariate

@@

# Sim D10: 10v10 Weak Covariate

@@

# Uninformative covariate

Here we repeat the previous four sections, using an uninformative covariate.

## Sim D5: DE Comparison 5v5

@@

## Sim D10: DE Comparison 10v10

@@

## Sim D5: 5v5 Weak Covariate

Here we'll repeat the above, but using a weaker informative covariate (that has noise
added to it).
This will be done for 100 random splits.

```{r d5-uc-uw, results='hide', message=FALSE}
rseed <- 445
sampleSize <- 5
nDE <- 500

if (!file.exists(resfile_d5_uninfCov_w)){
  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd,
                  uninformativeCovariate = TRUE, 
                  strongCovariate = FALSE,
                  mc.cores=nCores)
  saveRDS(de5, file = resfile_d5_uninfCov_w)
}else{
  de5 <- readRDS(file = resfile_d5_uninfCov_w)
}
```

## Sim D10: 10v10 Weak Covariate

```{r d10-uc-uw, results='hide', message=FALSE}
rseed <- 868
sampleSize <- 10
nDE <- 500

if(!file.exists(resfile_d10_uninfCov_w)){
  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, 
                   strongCovariate = FALSE,
                   mc.cores=nCores)
  saveRDS(de10, file = resfile_d10_uninfCov_w)
}else{
  de10 <- readRDS(file = resfile_d10_uninfCov_w)
}
```

@@

```{r, fig.width=18, fig.height=12}
plotMethodRanks(c(resfile_d5, resfile_d5_w, resfile_d10, resfile_d10_w,
                  resfile_d5_uninfCov, resfile_d5_uninfCov_w, resfile_d10_uninfCov, resfile_d10_uninfCov_w), 
                colLabels = c("5", "5 UW", "10", "10 UW",
                              "5 UI", "5 UI UW", "10 UI", "10 UI UW"), 
                alpha = 0.10, xlab = "Comparison", 
                excludeMethods = NULL)
```

>>>>

# results files that will be generated
# strong covariate
resfile_d5 <- file.path(resdir, "yeastIIH-results-de5.rds")
resfile_d10 <- file.path(resdir, "yeastIIH-results-de10.rds")

# weak covariate
resfile_d5_w <- file.path(resdir, "yeastIIHW-results-de5.rds")
resfile_d10_w <- file.path(resdir, "yeastIIHW-results-de10.rds")

# uninformative covariate
resfile_d5_uninfCov <- file.path(resdir, "yeastIIH-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "yeastIIH-results-de10-uninfCov.rds")

@@

  covar <- 1 / (1+exp(-runif(nrow(dds_full), 0, 10) + 5))

@@

  if(nDE > 0){
    # pick random set of nDE genes to add signal to
    DE <- sample(1:nrow(dds_test), nDE, prob = covar)
    truth[DE] <- TRUE

@@

  if (uninformativeCovariate){
    geneExp <- mutate(geneExp, ind_covariate = runif(length(covar)))
  }else if(!strongCovariate){
    geneExp <- mutate(geneExp, ind_covariate = pmin(1, abs(covar + rnorm(length(covar), 0, 0.25))))
  }

@@

```
# Non-null Comparisons with Strongly Informative Covariate

## D5S: DE 5v5 Strong

@@

@@

@@

# Non-null Comparisons with Strongly Informative Covariate

## D10S: DE 10v10 Strong

@@

```
# Non-null Comparisons with Weakly Informative Covariate

## D5W: DE 5v5 Weak

@@

## D10W: DE 10v10 Weak

@@

# Non-null Comparisons with Uninformative Covariate

Here we repeat the previous sections, using an uninformative covariate.

## D5U: DE 5v5 Uninformative

@@

## D10U: DE 10v10 Uninformative

@@

@@

```{r, fig.width=18, fig.height=12}
plotMethodRanks(c(resfile_d5, resfile_d5_w, resfile_d10, resfile_d10_w,
                  resfile_d5_uninfCov, resfile_d10_uninfCov), 
                colLabels = c("5 S", "5 W", "10 S", "10 W",
                              "5 UI", "10 UI"), 
                alpha = 0.10, xlab = "Comparison", 
                excludeMethods = NULL)
```

### 2018-09-13 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Refine informative covariate in yeast
Make strongly informative covariate stronger and weakly informative
covariate weaker.

# results files that will be generated
resfile_d5 <- file.path(resdir, "yeastII-results-de5.rds")
resfile_d10 <- file.path(resdir, "yeastII-results-de10.rds")
resfile_d5_w <- file.path(resdir, "yeastIIW-results-de5.rds")
resfile_d10_w <- file.path(resdir, "yeastIIW-results-de10.rds")

resfile_d5_uninfCov <- file.path(resdir, "yeastII-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "yeastII-results-de10-uninfCov.rds")
resfile_d5_uninfCov_w <- file.path(resdir, "yeastIIW-results-de5-uninfCov.rds")
resfile_d10_uninfCov_w <- file.path(resdir, "yeastIIW-results-de10-uninfCov.rds")

@@

  covar <- abs(rnorm(nrow(dds_full)))

@@

# Sim D5: DE Comparison 5v5

## Generate a list of SB results objects

@@

# Sim D10: DE Comparison 10v10

@@

# Sim D5: 5v5 Weak Covariate

@@

# Uninformative covariate

@@

## Sim D5: DE Comparison 5v5

@@

## Sim D5: 5v5 Weak Covariate

Here we'll repeat the above, but using a weaker informative covariate (that has noise
added to it).
This will be done for 100 random splits.

```{r d5-uc-uw, results='hide', message=FALSE}
rseed <- 445
sampleSize <- 5
nDE <- 2000

if (!file.exists(resfile_d5_uninfCov_w)){
  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd,
                  uninformativeCovariate = TRUE, 
                  strongCovariate = FALSE,
                  mc.cores=nCores)
  saveRDS(de5, file = resfile_d5_uninfCov_w)
}else{
  de5 <- readRDS(file = resfile_d5_uninfCov_w)
}
```

## Sim D10: 10v10 Weak Covariate

```{r d10-uc-uw, results='hide', message=FALSE}
rseed <- 868
sampleSize <- 10
nDE <- 2000

if(!file.exists(resfile_d10_uninfCov_w)){
  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, 
                   strongCovariate = FALSE,
                   mc.cores=nCores)
  saveRDS(de10, file = resfile_d10_uninfCov_w)
}else{
  de10 <- readRDS(file = resfile_d10_uninfCov_w)
}
```

>>>>

# results files that will be generated
# strong covariate
resfile_d5 <- file.path(resdir, "yeastII-results-de5.rds")
resfile_d10 <- file.path(resdir, "yeastII-results-de10.rds")

# weak covariate
resfile_d5_w <- file.path(resdir, "yeastIIW-results-de5.rds")
resfile_d10_w <- file.path(resdir, "yeastIIW-results-de10.rds")

# uninformative covariate
resfile_d5_uninfCov <- file.path(resdir, "yeastII-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "yeastII-results-de10-uninfCov.rds")

@@

  covar <- 1 / (1+exp(-runif(nrow(dds_full), 0, 10) + 5))

@@

# Non-null Comparisons with Strongly Informative Covariate

## D5S: DE 5v5 Strong

Here we'll carry out a DE comparison 5 versus 5 Snf2 samples,
where the groups are selected randomly and 2000 DE genes are added, and a
strongly informative covariate is used.
This will be done for 100 random splits. Note that we do not need to run null
comparisons, since these are done in `yeast-simulation.Rmd`, and the only
difference here is that we use a different under the alternative.

@@

## D10S: DE 10v10 Strong

@@

# Non-null Comparisons with Strongly Informative Covariate

## D5S: DE 5v5 Weak

@@

# Non-null Comparisons with Uninformative Covariate

@@

## D5U: DE 5v5 Uninformative

@@

### 2018-09-13 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
Refine informative covariate in yeast
Make strongly informative covariate stronger and weakly informative
covariate weaker.

# results files that will be generated
resfile_d5 <- file.path(resdir, "yeastH-results-de5.rds")
resfile_d10 <- file.path(resdir, "yeastH-results-de10.rds")
resfile_d5_w <- file.path(resdir, "yeastHW-results-de5.rds")
resfile_d10_w <- file.path(resdir, "yeastHW-results-de10.rds")

resfile_d5_uninfCov <- file.path(resdir, "yeastH-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "yeastH-results-de10-uninfCov.rds")
resfile_d5_uninfCov_w <- file.path(resdir, "yeastHW-results-de5-uninfCov.rds")
resfile_d10_uninfCov_w <- file.path(resdir, "yeastHW-results-de10-uninfCov.rds")

@@

# Sim D5: DE Comparison 5v5

@@

# Sim D5: 5v5 Weak Covariate

@@

## Sim D5: 5v5 Weak Covariate

Here we'll repeat the above, but using a weaker informative covariate (that has noise
added to it).
This will be done for 100 random splits.

```{r d5-uc-uw, results='hide', message=FALSE}
rseed <- 445
sampleSize <- 5
nDE <- 500

if(!file.exists(resfile_d5_uninfCov_w)){
  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd,
                  uninformativeCovariate = TRUE, 
                  strongCovariate=FALSE,
                  mc.cores=nCores)
  saveRDS(de5, file = resfile_d5_uninfCov_w)
}else{
  de5 <- readRDS(file = resfile_d5_uninfCov_w)
}
```

if(numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)

## Sim D10: 10v10 Weak Covariate

```{r d10-uc-uw, results='hide', message=FALSE}
rseed <- 868
sampleSize <- 10
nDE <- 500

if(!file.exists(resfile_d10_uninfCov_w)){
  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, 
                   strongCovariate = FALSE,
                   mc.cores=nCores)
  saveRDS(de10, file = resfile_d10_uninfCov_w)
}else{
  de10 <- readRDS(file = resfile_d10_uninfCov_w)
}
```

>>>>

# results files that will be generated
# strong covariate
resfile_d5 <- file.path(resdir, "yeastH-results-de5.rds")
resfile_d10 <- file.path(resdir, "yeastH-results-de10.rds")

# weak covariate
resfile_d5_w <- file.path(resdir, "yeastHW-results-de5.rds")
resfile_d10_w <- file.path(resdir, "yeastHW-results-de10.rds")

# uninformative covariate
resfile_d5_uninfCov <- file.path(resdir, "yeastH-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "yeastH-results-de10-uninfCov.rds")

@@

# Non-null Comparisons with Strongly Informative Covariate

## D5S: DE 5v5 Strong

@@

# Non-null Comparisons with Weakly Informative Covariate

## D5W: DE 5v5 Weak

@@

### 2018-09-13 datasets/RNA-Seq/yeast-simulation.Rmd
Refine informative covariate in yeast
Make strongly informative covariate stronger and weakly informative
covariate weaker.

# results files that will be generated
resfile_n5 <- file.path(resdir, "yeast-results-null5.rds")
resfile_n10 <- file.path(resdir, "yeast-results-null10.rds")
resfile_d5 <- file.path(resdir, "yeast-results-de5.rds")
resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")
resfile_d5_w <- file.path(resdir, "yeastW-results-de5.rds")
resfile_d10_w <- file.path(resdir, "yeastW-results-de10.rds")

resfile_n5_uninfCov <- file.path(resdir, "yeast-results-null5-uninfCov.rds")
resfile_n10_uninfCov <- file.path(resdir, "yeast-results-null10-uninfCov.rds")
resfile_d5_uninfCov <- file.path(resdir, "yeast-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "yeast-results-de10-uninfCov.rds")
resfile_d5_uninfCov_w <- file.path(resdir, "yeastW-results-de5-uninfCov.rds")
resfile_d10_uninfCov_w <- file.path(resdir, "yeastW-results-de10-uninfCov.rds")

@@

# Sim N5: Null Comparison 5v5

@@

# Sim N10: Null Comparison 10v10

Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
where the groups are selected randomly. This will be done for 100 random splits.

## Generate a list of SB results objects

@@

# Sim D5: DE Comparison 5v5

## Generate a list of SB results objects

@@

# Sim D10: DE Comparison 10v10

## Generate a list of SB results objects

@@

# Sim D5: 5v5 Weak Covariate

## Generate a list of SB results objects

@@

# Uninformative covariate

Here we repeat the previous four sections, using an uninformative covariate.

## Sim N5: Null Comparison 5v5

```{r n5-uc, results='hide', message=FALSE}
rseed <- 225
sampleSize <- 5
nDE <- 0

if (!file.exists(resfile_n5_uninfCov)){
  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, mc.cores = nCores)
  saveRDS(null5, file = resfile_n5_uninfCov)
}else{
  null5 <- readRDS(file = resfile_n5_uninfCov)
}
```

## Sim N10: Null Comparison 10v10

Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
where the groups are selected randomly. This will be done for 100 random splits.

```{r n10-uc, results='hide', message=FALSE}
rseed <- 837
sampleSize <- 10
nDE <- 0

if (!file.exists(resfile_n10_uninfCov)){
  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, mc.cores=nCores)
  saveRDS(null10, file = resfile_n10_uninfCov)
}else{
  null10 <- readRDS(file = resfile_n10_uninfCov)
}
```

## Sim D5: DE Comparison 5v5

@@

## Sim D5: 5v5 Weak Covariate

Here we'll repeat the above, but using a weaker informative covariate (that has noise
added to it).
This will be done for 100 random splits.

```{r d5-uc-uw, results='hide', message=FALSE}
rseed <- 445
sampleSize <- 5
nDE <- 2000

if (!file.exists(resfile_d5_uninfCov_w)){
  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd,
                  uninformativeCovariate = TRUE, 
                  strongCovariate = FALSE,
                  mc.cores=nCores)
  saveRDS(de5, file = resfile_d5_uninfCov_w)
}else{
  de5 <- readRDS(file = resfile_d5_uninfCov_w)
}
```

## Sim D10: 10v10 Weak Covariate

```{r d10-uc-uw, results='hide', message=FALSE}
rseed <- 868
sampleSize <- 10
nDE <- 2000

if(!file.exists(resfile_d10_uninfCov_w)){
  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE, 
                   weakCovariate = FALSE,
                   mc.cores=nCores)
  saveRDS(de10, file = resfile_d10_uninfCov_w)
}else{
  de10 <- readRDS(file = resfile_d10_uninfCov_w)
}
```

>>>>

# results files that will be generated
# null
resfile_n5 <- file.path(resdir, "yeast-results-null5.rds")
resfile_n10 <- file.path(resdir, "yeast-results-null10.rds")

# strong covariate
resfile_d5 <- file.path(resdir, "yeast-results-de5.rds")
resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")

# weak covariate
resfile_d5_w <- file.path(resdir, "yeastW-results-de5.rds")
resfile_d10_w <- file.path(resdir, "yeastW-results-de10.rds")

# uninformative covariate
resfile_d5_uninfCov <- file.path(resdir, "yeast-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "yeast-results-de10-uninfCov.rds")

@@

# Null Comparisons

## N5: Null 5v5

@@

## N10: Null 10v10

Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
where the groups are selected randomly. This will be done for 100 random splits.
Note that here the covariate-aware methods use a covariate that by definition is
uninformative (since there are no non-null observations).

@@

# Non-null Comparisons with Strongly Informative Covariate

## D5S: DE 5v5 Strong

Here we'll repeat the above, but for a DE comparison 5 versus 5 Snf2 samples,
where the groups are selected randomly and 2000 DE genes are added. 
We'll use a strongly informative covariate.
This will be done for 100 random splits.

@@

## D10S: DE 10v10 Strong

Here we'll repeat the above, but for a DE comparison 10 versus 10 Snf2 samples,
where the groups are selected randomly and 2000 DE genes are added.
We'll use a strongly informative covariate.
This will be done for 100 random splits.

@@

# Non-null Comparisons with Weakly Informative Covariate

## D5W: DE 5v5 Weak

Here we'll repeat the previous section, but using a weaker 
informative covariate (that has noise added to it).

@@

# Non-null Comparisons with Uninformative Covariate

Here we repeat the previous section, using an uninformative covariate.

## D5U: DE 5v5 Uninformative

@@
